---
title: "Session 1"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```




## What data do I have? 

This is how you need categorical data before plotting

```{r dataCreation1, exercise=TRUE}
inventedDF <- data.frame(schoolType=c("Public", "Private", "Other"),
                         Count=c(200, 20, 80))
inventedDF

```


## Prepating ggplot2

Nothing, apparently, happens:

```{r baseGGplot, exercise=TRUE}
inventedDF <- data.frame(schoolType=c("Public", "Private", "Other"),
                         Count=c(200, 20, 80))

library(ggplot2)

base=ggplot(data=inventedDF, # name of data frame
            aes(x=schoolType,# variables to use in each object
                y=Count))
```

## Default plot

```{r barsdefault, exercise=TRUE}
inventedDF <- data.frame(schoolType=c("Public", "Private", "Other"),
                         Count=c(200, 20, 80))

library(ggplot2)

base=ggplot(data=inventedDF,aes(x=schoolType,y=Count))
base + geom_bar(stat = 'identity') # object BAR plots "y", as is.
```

## Default plus Title

```{r barsTitle, exercise=TRUE}
inventedDF <- data.frame(schoolType=c("Public", "Private", "Other"),
                         Count=c(200, 20, 80))
#
# all text for titles
Titles=list(Ti="Nice title",STi="sub is here",XTi="Type of school",YTi="Amount of schools",Sou="Source: imaginary")
#
library(ggplot2)
base=ggplot(data=inventedDF,aes(x=schoolType,y=Count))
bar1= base + geom_bar(stat = 'identity')
#
# using texts
bar1 + labs(title = Titles$Ti,subtitle = Titles$STi,
            xlab=Titles$XTi,ylab=Titles$YTi,
            caption = Titles$Sou)
```

## Default plus annotating

```{r barsAnnot, exercise=TRUE}
inventedDF=data.frame(schoolType=c("Public", "Private", "Other"),Count=c(200, 20,80))
Titles=list(Ti="Nice title",STi="sub is here",XTi="Type of school",YTi="Amount of schools",Sou="Source: imaginary")
#
#texts for annotation
annots=list(forBar="HELLO")
#
library(ggplot2)
base=ggplot(data=inventedDF,aes(x=schoolType,y=Count)) 
bar1= base + geom_bar(stat = 'identity')
bar1=bar1 + labs(title = Titles$Ti,subtitle = Titles$STi,
                 xlab=Titles$XTi,ylab=Titles$YTi,
                 caption = Titles$Sou)
#
#annotating
bar1 + annotate(geom = 'text',x = 2,y=30,label=annots$forBar)
```

## Default changes: background

```{r barsBackground, exercise=TRUE}
inventedDF=data.frame(schoolType=c("Public", "Private", "Other"),Count=c(200, 20,80))
Titles=list(Ti="Nice title",STi="sub is here",XTi="Type of school",YTi="Amount of schools",Sou="Source: imaginary")
annots=list(forBar="HELLO")

library(ggplot2)
base= ggplot(data=inventedDF,aes(x=schoolType,y=Count)) 
#
# background
base= base + theme_classic()
#
bar1= base + geom_bar(stat = 'identity')
bar1= bar1 + labs(title = Titles$Ti,subtitle = Titles$STi,
                  xlab=Titles$XTi,ylab=Titles$YTi,
                  caption = Titles$Sou)
bar1 + annotate(geom = 'text',x = 2,y=30,label=annots$forBar)
```

## Default changes: position

```{r barsPosition, exercise=TRUE}
inventedDF=data.frame(schoolType=c("Public", "Private", "Other"),Count=c(200, 20,80))
Titles=list(Ti="Nice title",STi="sub is here",XTi="Type of school",YTi="Amount of schools",Sou="Source: imaginary")
annots=list(forBar="HELLO")
#
# label position on horizontal
NEWorder=c('Private','Other','Public')
#
library(ggplot2)
base= ggplot(data=inventedDF,aes(x=schoolType,y=Count)) 
base= base + theme_classic()

bar1= base + geom_bar(stat = 'identity')
bar1= bar1 + labs(title = Titles$Ti,subtitle = Titles$STi,
                  xlab=Titles$XTi,ylab=Titles$YTi,
                  caption = Titles$Sou)
# new position for annotation
bar1= bar1 + annotate(geom = 'text',x = 1,y=30,label=annots$forBar)
#
# repositioning bars
bar1 + scale_x_discrete(limits=NEWorder)
```

## Default changes: titles alignment

```{r barsChangeTit, exercise=TRUE}
inventedDF=data.frame(schoolType=c("Public", "Private", "Other"),Count=c(200, 20,80))
Titles=list(Ti="Nice title",STi="sub is here",XTi="Type of school",YTi="Amount of schools",Sou="Source: imaginary")
annots=list(forBar="HELLO")
NEWorder=c('Private','Other','Public')

library(ggplot2)
base= ggplot(data=inventedDF,aes(x=schoolType,y=Count)) 
base= base + theme_classic()
bar1= base + geom_bar(stat = 'identity')
bar1= bar1 + labs(title = Titles$Ti,subtitle = Titles$STi,
                  xlab=Titles$XTi,ylab=Titles$YTi,
                  caption = Titles$Sou)
bar1= bar1 + annotate(geom = 'text',x = 1,y=30,label=annots$forBar)
bar1= bar1 + scale_x_discrete(limits=NEWorder)
# realignment of texts in titles
bar1 + theme(plot.title= element_text(hjust= 0.5),
             plot.subtitle= element_text(hjust= 0.5),
             plot.caption= element_text(hjust= 0))

```

## Default changes: color and legend

Run the code, then:

* Add in theme: **legend.position='top'**
* Add in scale_fill_brewer: **breaks=NEWorder**
* Make a change in guides: **guides(fill = "none")**

```{r barsChangeLegCol, exercise=TRUE}
inventedDF=data.frame(schoolType=c("Public", "Private", "Other"),Count=c(200, 20,80))
Titles=list(Ti="Nice title",STi="sub is here",XTi="Type of school",YTi="Amount of schools",Sou="Source: imaginary")
annots=list(forBar="HELLO")
NEWorder=c('Private','Other','Public')

library(ggplot2)
base= ggplot(data=inventedDF,aes(x=schoolType,y=Count)) 
base= base + theme_classic()
#
# Adding variable for coloring
bar1= base + geom_bar(stat = 'identity', aes(fill=schoolType)) 
#
bar1= bar1 + labs(title = Titles$Ti,subtitle = Titles$STi,
                  xlab=Titles$XTi,ylab=Titles$YTi,
                  caption = Titles$Sou)
bar1= bar1 + annotate(geom = 'text',x = 1,y=30,label=annots$forBar)
bar1= bar1 + scale_x_discrete(limits=NEWorder)
bar1= bar1 + theme(plot.title= element_text(hjust= 0.5),
                   plot.subtitle= element_text(hjust= 0.5),
                   plot.caption= element_text(hjust= 0))
#
# color and legend changes
bar1 + scale_fill_brewer(palette = "Accent") + guides(fill = "legend")

```
