---
title: "Session 1"
output: learnr::tutorial
runtime: shiny_prerendered
---

```{r setup, include=FALSE}
library(learnr)
knitr::opts_chunk$set(echo = FALSE)
```




## What data do I have? 

In general, it all starts with a data file. In my case, I have the data in this link:
```{r dataLink, exercise=FALSE,eval=TRUE, echo=TRUE}
location="https://github.com/EvansDataScience/data/raw/master/"
file="hsb_ok.sav"

linkToFile=paste0(location,file)
```

The object **link** only has the location of the file, then we can read it like this:

```{r openFileexercise=FALSE,eval=TRUE, echo=TRUE}
library(foreign)

hsb=read.spss(linkToFile,to.data.frame = T)
```
The object **hsb** has tha data. Software like R can try to recognize the data types, but it may not always get it right:

```{r strHSB, exercise=TRUE,eval=TRUE, echo=TRUE}
str(hsb)
```

We need to make sure we get it right!

From the previous result: Which remain as factor, as character, or as ordinal? (check [here](https://github.com/UWDataScience2020/data/blob/master/hsb_ok_metadata.pdf))

```{r changeType, exercise=TRUE,eval=FALSE, echo=TRUE}
as.ordered(hsb$SES)
```
Data types are critical for selecting the right visual.

## Plot components

1. **BASE**

You may want to create a **base* component, just to inform the data frame you are using and the variable(s) to plot:
```{r baseGGplot, eval=TRUE, echo=TRUE}
library(ggplot2)

base=ggplot(data=hsb, # data frame
            aes(x=MATH)) # what to plot
```

Notice that in **aes**thetics you tell what variable will be used to represent some information in the plot:

* Dimensions x, y, z .
* Color.
* Size.
* Label.

You may need to customize the **aes** in other elements.

2. **GEOMs**

Let's make a simple bar plot:

```{r barsdefault, exercise=TRUE}
#
plot1= base + geom_histogram()
# I need this to see the result:
plot1
```

The GEOM selected depends on the data type.


3. **TITLE**

```{r plot1Title, exercise=TRUE, eval=TRUE}
# all text for titles
Titles=list(Title="Nice title",
            SubTi="sub is here",
            XTi="Level of MATH",
            YTi="Amount of students",
            Sou="Source: imaginary")
#
# using the previous texts
plot2= plot1 + labs(title = Titles$Title,
                    subtitle = Titles$SubTi,
                    x=Titles$XTi,
                    y=Titles$YTi,
                    caption = Titles$Sou)
plot2
```

4. **ANNOTATION**

```{r plot1Annot, exercise=TRUE, eval=TRUE}

#texts for annotation
message="HELLO"
#
#annotating
plot3=plot2 + annotate(geom = 'text',
                       x =2,y=30, #position
                       label=message)
plot3
```

5. **BACKGROUND**


```{r barsBackground, exercise=TRUE}
plot4=plot3+ theme_classic()
plot4
```

## Default changes: position

```{r barsPosition, exercise=TRUE}
inventedDF=data.frame(schoolType=c("Public", "Private", "Other"),Count=c(200, 20,80))
Titles=list(Ti="Nice title",STi="sub is here",XTi="Type of school",YTi="Amount of schools",Sou="Source: imaginary")
annots=list(forBar="HELLO")
#
# label position on horizontal
NEWorder=c('Private','Other','Public')
#
library(ggplot2)
base= ggplot(data=inventedDF,aes(x=schoolType,y=Count)) 
base= base + theme_classic()

bar1= base + geom_bar(stat = 'identity')
bar1= bar1 + labs(title = Titles$Ti,subtitle = Titles$STi,
                  xlab=Titles$XTi,ylab=Titles$YTi,
                  caption = Titles$Sou)
# new position for annotation
bar1= bar1 + annotate(geom = 'text',x = 1,y=30,label=annots$forBar)
#
# repositioning bars
bar1 + scale_x_discrete(limits=NEWorder)
```

## Default changes: titles alignment

```{r barsChangeTit, exercise=TRUE}
inventedDF=data.frame(schoolType=c("Public", "Private", "Other"),Count=c(200, 20,80))
Titles=list(Ti="Nice title",STi="sub is here",XTi="Type of school",YTi="Amount of schools",Sou="Source: imaginary")
annots=list(forBar="HELLO")
NEWorder=c('Private','Other','Public')

library(ggplot2)
base= ggplot(data=inventedDF,aes(x=schoolType,y=Count)) 
base= base + theme_classic()
bar1= base + geom_bar(stat = 'identity')
bar1= bar1 + labs(title = Titles$Ti,subtitle = Titles$STi,
                  xlab=Titles$XTi,ylab=Titles$YTi,
                  caption = Titles$Sou)
bar1= bar1 + annotate(geom = 'text',x = 1,y=30,label=annots$forBar)
bar1= bar1 + scale_x_discrete(limits=NEWorder)
# realignment of texts in titles
bar1 + theme(plot.title= element_text(hjust= 0.5),
             plot.subtitle= element_text(hjust= 0.5),
             plot.caption= element_text(hjust= 0))

```

## Default changes: color and legend

Run the code, then:

* Add in theme: **legend.position='top'**
* Add in scale_fill_brewer: **breaks=NEWorder**
* Make a change in guides: **guides(fill = "none")**

```{r barsChangeLegCol, exercise=TRUE}
inventedDF=data.frame(schoolType=c("Public", "Private", "Other"),Count=c(200, 20,80))
Titles=list(Ti="Nice title",STi="sub is here",XTi="Type of school",YTi="Amount of schools",Sou="Source: imaginary")
annots=list(forBar="HELLO")
NEWorder=c('Private','Other','Public')

library(ggplot2)
base= ggplot(data=inventedDF,aes(x=schoolType,y=Count)) 
base= base + theme_classic()
#
# Adding variable for coloring
bar1= base + geom_bar(stat = 'identity', aes(fill=schoolType)) 
#
bar1= bar1 + labs(title = Titles$Ti,subtitle = Titles$STi,
                  xlab=Titles$XTi,ylab=Titles$YTi,
                  caption = Titles$Sou)
bar1= bar1 + annotate(geom = 'text',x = 1,y=30,label=annots$forBar)
bar1= bar1 + scale_x_discrete(limits=NEWorder)
bar1= bar1 + theme(plot.title= element_text(hjust= 0.5),
                   plot.subtitle= element_text(hjust= 0.5),
                   plot.caption= element_text(hjust= 0))
#
# color and legend changes
bar1 + scale_fill_brewer(palette = "Accent") + guides(fill = "legend")

```
